---
author: "Michael Koohafkan"
title: "How to use RAStestR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use RAStestR} 
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The R package **RAStestR** provides a library of functions for reading,
manipulating and visualizing HEC-RAS output data. Specifically, it 
accesses the HDF5 data file associated with an HEC-RAS "plan". This
vignette provides a brief introduction to the basic functionality of
**RAStestR**.


## An example project

Outputs from a simple quasi-unsteady sediment transport simulation are included with 
**RAStestR**. Basic information about the plan can be displayed with 
`list_plan_info`.

```{r}
library(RAStestR)
ras.file = system.file("sample-data/SampleQuasiUnsteady.hdf", package = "RAStestR")
list_plan_info(ras.file)
```

**RAStestR** provides functions for accessing data on model geometry and simulation outputs.
Simulation outputs are read into R via two basic functions: `read_standard()` and
`read_sediment()`. Most hydraulic data such as flow, water surface elevation, shear stress,
etc. are contained in single tables which are read in using `read_standard()`. Some outputs
from sediment transport models reported by grain class in individual tables; these tables
can be read in and automatically grouped using `read_sediment()`. In the example below, the
standard table "Shear Stress" and the sediment table "Mass Bed Change Cum" are read into R.

```{r}
shear.stress = read_standard(ras.file, "Shear Stress")
vol.change.cum = read_sediment(ras.file, "Vol Bed Change Cum")
```

You can also choose to extract data for only a subset of output times, river stations, or
grain classes. **RAStestR** provides listing functions to retrieve possible values.

```{r}
list_output_times(ras.file)
list_output_times(ras.file)
list_grain_classes(ras.file)
```

`read_standard` and `read_sediment` return wide-format tables which match the row-column
structure of the HDF file. extracting subsets of the data is more convenient with wide
tables as rows correspond to output times and columns correspond to river stations.
However, many operations in R are more convenient to use with
long-format tables. The `to_longtable` function collapses the cross-section columns in
a wide table automatically. The `to_widetable()` function can convert a long table back 
into the original wide format. The R package **dplyr** provides a rich environment for
manipulating long-format tables.

```{r}
# wide table format (first five columns)
head(shear.stress[1:5])
# long table format
shear.stress.long = to_longtable(shear.stress, "ShearStress")
head(shear.stress.long)
# convert from long-format to wide-format
#to_widetable(shear.stress.long, "Station", "ShearStress")
```

`read_standard`, `read_sediment` and `to_longtable` read the river stations, time stamps 
and grain classes as characters. This is inconvenient for e.g. plotting results over time
or along a reach. The function `reformat_fields()` will convert the time stamps
to `POSIXct` datetime format, and river stations to `numeric` format.

```{r}
shear.stress.long.fmt = reformat_fields(shear.stress.long)
```

